---
title: "sz_prs_icd10_analysis"
author: "Carla Arx칠"
date: "2025-05-21"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# INTRODUCTION TO SCRIPT
SCRIPT PURPOSE: This script investigates the relationship between the polygenic risk score for schizophrenia (SCZ-PRS) and a wide range of medical conditions as coded in ICD-10, using data from the UK Biobank.

# GENERAL GOAL:
Evaluate how genetic risk for schizophrenia relates to health outcomes in the general population, both in individuals with mental health diagnoses (ICD-10 Chapter F) and those without.

Specifically, this script:
 1. Merges SCZ-PRS residuals with ICD-10 diagnostic data from the UK Biobank.
 2. Constructs two datasets:
    a. Full dataset including individuals with ICD-10 Chapter F diagnoses.
    b. Filtered dataset excluding individuals with mental diagnoses (Chapter V).
 3. Analyzes the relationship between SCZ-PRS and disease prevalence:
    a. Using individual ICD-10 codes and grouped ICD-10 categories.
    b. Applying regression models (linear, log, exponential, power) and quantile analysis.
 4. Computes statistical and descriptive indicators

# OVERVIEW OF APPROACHES:

1. Load both files (PRS.PC1.residual and diagnostics). 
2. Transform the diagnoses (X41270.0.*) to long format. 
3. Extract the first 3 characters of the ICD-10 codes. 
4. Merge the data with the PRS only for the individuals who have it. 
5. Create two datasets: 
- One complete 
- One excluding patients from chapter V
6. Generate a binary diagnosis matrix by code. 
7. Create regression models by code with different types of fitting (linear, log, exponential). 
8. Filter only codes with at least 100 cases. 
9. Correct for multiple comparisons (FDR). 
10. Prevalence study by 25 quantiles of PRS.

```{r}
# ========================
# ==== LOAD LIBRARIES ====
# ========================
suppressMessages(library(dplyr))
suppressMessages(library(tidyr))
suppressMessages(library(stringr))
suppressMessages(library(purrr))
suppressMessages(library(broom))
suppressMessages(library(ggplot2))
suppressMessages(library(scales))
suppressMessages(library(tidyverse))
suppressMessages(library(gridExtra))
suppressMessages(library(knitr))
suppressMessages(library(kableExtra))
suppressMessages(library(grid))
suppressMessages(library(ggpubr))
suppressMessages(library(pheatmap))
suppressMessages(library(RColorBrewer))
```

```{r}
rm(list=ls(all.names=TRUE))
invisible(gc())
```

```{r}
# ===================
# ==== LOAD DATA ====
# ===================
df_prs <- read.table("3_PRS_PC1_res.txt", header = TRUE) 
df_icd <- read.table("ukb668761_41270_wb.txt", 
                     header = TRUE, 
                     sep = "\t",                   
                     fill = TRUE,
                     quote = "",
                     stringsAsFactors = FALSE,
                     check.names = FALSE)    
```

```{r}
### 2. Transform diagnostic data to long format
df_icd_long <- df_icd %>%
  select(-FID) %>%
  pivot_longer(cols = starts_with("X41270"), names_to = "field", values_to = "code") %>%
  filter(!is.na(code), code!="") %>%
  mutate(code3 = substr(code, 1, 3 )) %>%
  distinct(IID, code3)

# Number of unique values in the column code3
length(unique(df_icd_long$code3))
```

```{r}
### 3. Create a binary matrix (one column for each 3-digit ICD-10 code)
df_icd_bin <- df_icd_long %>%
  mutate(present = 1) %>%
  pivot_wider(names_from = code3, values_from = present, values_fill = 0)
```

```{r}
### 4. Join with PRS
df_all <- df_prs %>%
  left_join(df_icd_bin, by = "IID")
```

```{r}
### 5. Separate datasets 
# Complete dataset
df_with_F <- df_all

# 1. Detecta todas las columnas que comienzan con "F"
f_cols <- grep("^F", names(df_all), value = TRUE)
f_cols <- setdiff(f_cols, "FID")

# Dataset excluding F (chapter V)
df_no_F <- df_all %>%
  mutate(across(all_of(f_cols), ~ as.numeric(as.character(.x)))) %>%
  filter(rowSums(select(., all_of(f_cols)), na.rm = TRUE) == 0) %>%
  select(-all_of(f_cols))

# Get list of codes
icd_codes <- setdiff(names(df_all), c("FID", "IID", "PRS.PC1.residual"))
```

```{r}
code_counts <- df_no_F %>%
  select(-FID, -IID, -PRS.PC1.residual) %>%
  summarise(across(everything(), ~ sum(.x, na.rm = TRUE))) %>%
  pivot_longer(cols = everything(), names_to = "code", values_to = "n_individuals")

codes <- code_counts %>%
  filter(n_individuals >= 100) %>%
  pull(code)

df_no_F_filtered <- df_no_F %>%
  select(FID, IID, PRS.PC1.residual, all_of(codes))
```

```{r}
### 8. Save results
#write.csv(df_no_F_filtered, "df_no_F_filtered.csv", row.names = FALSE)
#df_no_F_filtered <- read.csv("df_no_F_filtered.csv", stringsAsFactors = FALSE)
```

```{r}
#5.1 check if it follows a normal distribution
hist(df_no_F$PRS.PC1.residual, breaks = 30, main = "PRS Histogram", xlab = "PRS")
```
```{r}
qqnorm(df_no_F$PRS.PC1.residual)
qqline(df_no_F$PRS.PC1.residual, col = "red")
```
# MODEL ADJUST
Objetivo:
Evaluar, para cada diagn칩stico (code), qu칠 modelo (lineal, logar칤tmico, exponencial) se ajusta mejor.
Comparar residuos para seleccionar el mejor modelo.
Visualizarlo con un Manhattan plot por modelo, donde cada punto represente un diagn칩stico (code) y el eje Y sea, por ejemplo:
el -log10(p) de ese modelo (si pruebas significancia),
o la R (explicaci칩n de varianza),
o el residual sum of squares (RSS) (para evaluar error).

```{r}
# A침adir columna de cuantiles a tu dataframe
df_filtered <- df_no_F_filtered %>%
  mutate(PRS_quantile = ntile(PRS.PC1.residual, 25))
```

```{r}
# Identificar las columnas de diagn칩stico: todas excepto ID y PRS
diagnosis_cols <- setdiff(colnames(df_filtered), c("FID", "IID", "PRS.PC1.residual", "PRS_quantile"))

# Pivot to long format
df_long <- df_filtered %>%
  pivot_longer(
    cols = all_of(diagnosis_cols),
    names_to = "code",
    values_to = "diagnosis"
  ) %>%
  select(FID, IID, PRS_quantile, code, diagnosis) 

# Calcular prevalencia y n칰mero de personas 칰nicas por c칩digo y cuantil
prevalence_by_quantile <- df_long %>%
  group_by(code, PRS_quantile) %>%
  summarise(
    prevalence = mean(diagnosis, na.rm = TRUE),
    n = n_distinct(IID),
    cases = sum(diagnosis == 1, na.rm = TRUE),  # <-- n칰mero de diagn칩sticos positivos
    .groups = "drop"
  )

# Verifica resultado
head(prevalence_by_quantile)
```

```{r}
# Renombrar para usar PRS en lugar de PRS_quantile
prevalence_by_quantile <- prevalence_by_quantile %>%
  rename(PRS = PRS_quantile)
```

```{r}
#Identify columns with diagnosis F
f_cols <- grep("^F", names(df_all), value = TRUE)

# Seleccionar solo las columnas relevantes: excluir ID, IID, PRS
cols_to_use <- setdiff(f_cols, c("FID", "IID", "PRS_qtile", "PRS.PC1.residual"))

df_all_F <- df_all %>%
  mutate(any_F = as.integer(rowSums(select(., all_of(cols_to_use)), na.rm = TRUE) > 0))

code_interest <- "any_F"

# 4. Plot
# Calcular prevalence per quantils
prevalence_plot_all_F <- df_all_F %>%
  group_by(PRS_qtile) %>%
  summarise(
    N = n(),
    n_affected = sum(.data[[code_interest]], na.rm = TRUE),
    prevalence = n_affected / N
  ) %>%
  ggplot(aes(x = PRS_qtile, y = prevalence, group = 1)) +
  geom_point(aes(color = prevalence), size = 2) +                                
  geom_text(
    data = . %>% filter(PRS_qtile %in% c(1, 25)),
    aes(label = paste0(round(prevalence * 100, 1), "%")),
    vjust = -1.2, size = 3.5
  ) +
  geom_smooth(method = "gam", formula = y ~ s(x, bs = "cs"), 
              aes(group = 1), color = "steelblue", se = TRUE) +                 # Curva suavizada con IC
  scale_color_viridis_c(option = "viridis", direction = -1, name = "Prevalence") +
  scale_x_discrete(drop = FALSE) +                                              # Mostrar todos los cuantiles
  labs(
    title = paste("Prevalence of", "Mental and behavioural disorders", "by PRS"),
    x = "PRS Quantile",
    y = "Prevalence"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(face = "bold", size = 12),
    axis.text = element_text(size = 10),
    axis.title = element_text(size = 10),
    legend.text = element_text(size = 10),
    legend.title = element_text(size = 10)
  )

# 5. Mostrar el gr치fico
prevalence_plot_all_F
```

#Prior ASSOCIATION
```{r}
### 6. Function for code analysis
analyze_icd <- function(df, prs_col = "PRS.PC1.residual", min_n = 100) {
  results <- map_dfr(icd_codes, function(code) {
    n_case <- sum(df[[code]], na.rm = TRUE)
    if (n_case < min_n) return(NULL)
    
    model_df <- df %>%
      select(PRS = all_of(prs_col), diagnosis = all_of(code)) %>%
      filter(!is.na(PRS))
    
    # Model
    log_glm <- glm(diagnosis ~ log(PRS - min(PRS) + 1e-5), family = "binomial", data = model_df)
    
    tibble(
      code = code,
      n_cases = n_case,
      beta = coef(log_glm)[2],
      p_value = summary(log_glm)$coefficients[2, 4],
      OR = exp(coef(log_glm)[2]),
      SE = summary(log_glm)$coefficients[2, "Std. Error"]
    )
  })
  
  # Correcction by FDR
  results <- results %>%
    mutate(
      p_value_adj = p.adjust(p_value, method = "fdr")
    )
  return(results)
}
```
 
```{r}
### 7. Execute analysis
results_with_F <- analyze_icd(df_with_F)
results_no_F <- analyze_icd(df_no_F)
```

```{r}
### 8. Save results
#write.csv(results_with_F, "results_with_F.csv", row.names = FALSE)
#write.csv(results_no_F, "results_no_F.csv", row.names = FALSE)
```

```{r}
#results_with_F <- read.csv("results_with_F.csv", stringsAsFactors = FALSE)
#results_no_F <- read.csv("results_no_F.csv", stringsAsFactors = FALSE)
```

```{r}
results_ordered_F <- results_no_F %>%
  arrange(p_value_adj)
```

```{r}
results_no_F_filter <- results_no_F %>%
  dplyr::filter(p_value_adj < 0.05)
```

```{r}
#False discovery rate
# ---------------------
# 2. Preparar columnas necesarias
# ---------------------
# Asigna letra del c칩digo ICD-10
results_no_F <- results_no_F %>%
  mutate(
    chapter = str_sub(code, 1, 1),
    chapter_name = case_when(
      chapter == "A" | chapter == "B" ~ "Infecntious and parasitic",
      chapter == "C" | chapter == "D" & as.numeric(str_sub(code, 2, 3)) <= 48 ~ "Neoplasms",
      chapter == "D" ~ "Blood and hematopoietic organs",
      chapter == "E" ~ "Endocrine, nutritional and metabolic",
      chapter == "F" ~ "Mental and behavioural disorders",
      chapter == "G" ~ "Nervous system",
      chapter == "H" & as.numeric(str_sub(code, 2, 3)) <= 59 ~ "Eye and adnexa",
      chapter == "H" ~ "Ear and mastoid process",
      chapter == "I" ~ "Circulatory system",
      chapter == "J" ~ "Respiratory system",
      chapter == "K" ~ "Digestive system",
      chapter == "L" ~ "Skin and subcutaneous tissue",
      chapter == "M" ~ "Musculoskeletal system and connective tissue",
      chapter == "N" ~ "Genitourinary system",
      chapter == "O" ~ "Pregnancy, childbirth and the puerperium",
      chapter == "P" ~ "Perinatal period",
      chapter == "Q" ~ "Congenital malformations",
      chapter == "R" ~ "Abnormal clinical and laboratory findings",
      chapter == "S" | chapter == "T" ~ "Injury and poisoning",
      chapter == "V" | chapter == "W" | chapter == "X" | chapter == "Y" ~ "External causes",
      chapter == "Z" ~ "Factors influencing health status",
      TRUE ~ "Codes for special purposes"
    )
  )

data <- results_no_F %>%
  mutate(
    neg_log10_p = -log10(p_value_adj),
    direction = ifelse(beta >= 0, "Positive", "Negative"),
    index = row_number()
  )

threshold <- -log10(0.05)

# 2. Crear una tabla con colores por cap칤tulo
chapters <- sort(unique(data$chapter))
color_palette <- scales::hue_pal()(length(chapters))  # paleta de colores autom치ticos
chapter_colors <- setNames(color_palette, chapters)   # nombrar por letra

data <- data %>%
  arrange(chapter_name) %>%
  mutate(index = row_number())

# ---------------------
# 3. Crear gr치fico
# ---------------------
# 游꿛 Gr치fico Manhattan con direcci칩n y etiquetado
ggplot(data, aes(x = index, y = neg_log10_p)) +
  geom_point(aes(color = chapter_name, shape = direction), size = 3) +
  geom_hline(yintercept = threshold, color = "red", linetype = "dashed") +
  geom_text(
    data = filter(data, neg_log10_p > threshold),
    aes(label = code),
    vjust = -1, size = 3
  ) +
  scale_shape_manual(values = c("Positive" = 24, "Negative" = 25)) +
  labs(
    title = "Manhattan Plot by ICD-10 (logistic model)",
    x = "Phenotype",
    y = expression(-log[10](p[adj])),
    color = "",
    shape = ""
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    legend.position = "bottom"
  )
```


```{r}
# Adjust models

# Funciones para obtener p-valores
get_lm_pvalue <- function(model) {
  fstat <- summary(model)$fstatistic
  if (is.null(fstat)) return(NA_real_)
  pf(fstat["value"], fstat["numdf"], fstat["dendf"], lower.tail = FALSE)
}

get_nls_b_pvalue <- function(model) {
  tryCatch(
    summary(model)$coefficients["b", "Pr(>|t|)"],
    error = function(e) NA_real_
  )
}

get_lm_r2 <- function(model) {
  summary(model)$r.squared
}

get_nls_r2 <- function(model, data) {
  if (is.null(model) || is.null(data)) return(NA_real_)
  prevalence <- as.numeric(data$prevalence)
  if (anyNA(prevalence)) return(NA_real_)  # Evita problemas con NAs
  ss_res <- sum(residuals(model)^2)
  ss_tot <- sum((prevalence - mean(prevalence))^2)
  1 - ss_res / ss_tot
}

safe_nls <- safely(function(data) {
  nls(prevalence ~ a * PRS^b, start = list(a = 1, b = 1), data = data)
})

# Models adjust
model_fits <- prevalence_by_quantile %>%
  group_by(code) %>%
  nest() %>%
  mutate(
    lin_model = map(data, ~ lm(prevalence ~ PRS, data = .x)),
    log_model = map(data, ~ lm(prevalence ~ log(PRS), data = .x)),
    pw_model_raw = map(data, ~ safe_nls(.x)),
    pw_model = map(pw_model_raw, "result"),
    
    # Nuevo modelo exponencial: log(prevalence) ~ PRS
    exp_model = map(data, ~ {
  data_clean <- .x %>% filter(prevalence > 0)
  if (nrow(data_clean) < 2) return(NULL)  # evitar modelos con datos insuficientes
  lm(log(prevalence) ~ PRS, data = data_clean)
}),

    aic_lin = map_dbl(lin_model, AIC),
    aic_log = map_dbl(log_model, AIC),
    aic_pw = map_dbl(pw_model, ~ if (!is.null(.x)) AIC(.x) else NA_real_),
    aic_exp = map_dbl(exp_model, AIC),

    rss_lin = map_dbl(lin_model, ~ sum(residuals(.x)^2)),
    rss_log = map_dbl(log_model, ~ sum(residuals(.x)^2)),
    rss_pw = map_dbl(pw_model, ~ if (!is.null(.x)) sum(residuals(.x)^2) else NA_real_),
    rss_exp = map_dbl(exp_model, ~ sum(residuals(.x)^2))  # sobre log(prevalence)
  )

model_fits <- model_fits %>%
  mutate(
    r2_lin = map_dbl(lin_model, get_lm_r2),
    r2_log = map_dbl(log_model, get_lm_r2),
    r2_pw  = map2_dbl(pw_model, data, get_nls_r2),
    r2_exp = map_dbl(exp_model, get_lm_r2)
  )

model_selection_with_p <- model_fits %>%
  transmute(
    code,
    best_model = pmap_chr(
      list(r2_lin, r2_log, r2_pw, r2_exp),
      ~ c("linear", "log", "power", "exp_lm")[which.max(c(...))]
    ),
    min_rss = pmap_dbl(list(rss_lin, rss_log, rss_pw, rss_exp), ~ min(c(...))),
    aic_diff = pmap_dbl(list(aic_lin, aic_log, aic_pw, aic_exp), ~ min(c(...)) - max(c(...))),

    p_value = pmap_dbl(
      list(best_model, lin_model, log_model, pw_model, exp_model),
      function(model_type, lin, log, pw, exp) {
        if (model_type == "linear")  return(get_lm_pvalue(lin))
        if (model_type == "log")     return(get_lm_pvalue(log))
        if (model_type == "power")   return(get_nls_b_pvalue(pw))
        if (model_type == "exp_lm")  return(get_lm_pvalue(exp))
      }
    ),

    r_squared = pmap_dbl(
  list(best_model, lin_model, log_model, pw_model, exp_model, data),
  function(model_type, lin, log, pw, exp, dat) {
    if (is.null(dat)) return(NA_real_)
    if (model_type == "linear") return(get_lm_r2(lin))
    if (model_type == "log")    return(get_lm_r2(log))
    if (model_type == "power")  return(get_nls_r2(pw, dat))
    if (model_type == "exp_lm")    return(get_lm_r2(exp))
    NA_real_
  }
    ),
  
  coef_estimate = pmap_dbl(
      list(best_model, lin_model, log_model, pw_model, exp_model),
      function(model_type, lin, log, pw, exp) {
        if (model_type == "linear") return(coef(lin)[["PRS"]])
        if (model_type == "log")    return(coef(log)[["log(PRS)"]])
        if (model_type == "power") {
          if (is.null(pw)) return(NA_real_)
          return(coef(pw)[["b"]])
        }
        if (model_type == "exp_lm") return(coef(exp)[["PRS"]])
        NA_real_
      }
    )
    ) %>%
  select(code, best_model, coef_estimate, min_rss, aic_diff, p_value, r_squared)


model_selection_with_p_adj <- model_selection_with_p %>%
  ungroup() %>%
  mutate(p_value_adj = p.adjust(p_value, method = "fdr"))

model_selection_adj <- model_selection_with_p_adj %>%
  filter(p_value_adj < 0.05)  %>%
  arrange(p_value_adj)  %>%
  select(
    1:5,
    7,
    6,
    8
  )

model_selection <- model_selection_adj %>%
  filter(r_squared > 0.45)

#write.csv(model_selection_adj, "table.csv", row.names = FALSE)
```

```{r}
#Model_selection table
model_selection <- model_selection %>%
  mutate(code_names = recode(code,
    "J43" = "J43: Emphysema",
    "K92" = "K92: Other digestive diseases",
    "M18" = "M18: Arthrosis of first carpometacarpal joint",
    "M65" = "M65: Synovitis and tenosynovitis",
    "M17" = "M17: Knee Osteoarthritis",
    "E11" = "E11: Type 2 Diabetes",
    "E66" = "E66: Obesity",
    "N17" = "N17: Acute Renal Failure",
    "G56" = "G56: Mononeuropathies of upper limb",
    "K59" = "K59: Other intestinal disorders",
    "Z37" = "Z37: Childbirth",
    "R10" = "R10: Abdominal Pain",
    "M23" = "M23: Internal derangement of knee",
    "M19" = "M19: Other arthrosis",
    "K62" = "K62: Anal and rectum disorders"
  ))
```

```{r}
#Heatmap
# Afegim columna de controls
prev_df <- prevalence_by_quantile %>%
  mutate(controls = n - cases)

prev_df <- prev_df %>%
  rename(PRS_quantile = PRS)

#filtrem per aquells que tenen el model significatiu
prev_df_filtered <- prev_df %>%
  semi_join(model_selection, by = "code")

# 1. Calcular OR (referencia Q13)
calculate_or_vs_q13_fast <- function(df, ref_quantile = 13) {
  df <- df %>%
    mutate(PRS = factor(PRS_quantile, levels = sort(unique(PRS_quantile)))) %>%
    mutate(PRS = relevel(PRS, ref = as.character(ref_quantile)))
  
  model <- glm(cbind(cases, controls) ~ PRS, data = df, family = binomial)
  
  broom::tidy(model, exponentiate = TRUE, conf.int = TRUE) %>%
    filter(term != "(Intercept)") %>%
    mutate(quantile = as.integer(gsub("PRS", "", term))) %>%
    select(quantile, estimate, p.value) %>%
    bind_rows(tibble(quantile = ref_quantile, estimate = 1, p.value = NA)) %>%
    arrange(quantile)
}

# Calculate OR vs Q13
or_by_code_q13 <- prev_df_filtered %>%
  group_by(code) %>%
  group_modify(~ calculate_or_vs_q13_fast(.x)) %>%
  ungroup()

# Prepare matrix for heatmap
df_or <- or_by_code_q13 %>%
  mutate(logOR = log(estimate))  # Optional: log-transform

or_matrix <- df_or %>%
  select(code, quantile, estimate) %>%
  pivot_wider(names_from = quantile, values_from = estimate) %>%
  column_to_rownames("code")

# Replace these mappings with your actual disease names
code_names <- c("J43" = "J43: Emphysema", 
                "K92" = "K92: Other digestive diseases", 
                "M18" = "M18: Arthrosis of first carpometacarpal joint", 
                "M65" = "M65: Synovitis and tenosynovitis", 
                "M17" = "M17: Knee Osteoarthritis", 
                "E11" = "E11: Type 2 Diabetes", 
                "E66" = "E66: Obesity", 
                "N17" = "N17: Acute Renal Failure", 
                "G56" = "G56: Mononeuropathies of upper limb", 
                "K59" = "K59: Other intestinal disorders", 
                "Z37" = "Z37: Childbirth", 
                "R10" = "R10: Abdominal Pain", 
                "M23" = "M23: Internal derangement of knee", 
                "M19" = "M19: Other arthrosis", 
                "K62" = "K62: Anal and rectum disorders")

rownames(or_matrix) <- ifelse(rownames(or_matrix) %in% names(code_names), 
                              code_names[rownames(or_matrix)], 
                              rownames(or_matrix))

# Define color palette centered at OR = 1
breaks <- seq(min(or_matrix, na.rm = TRUE), max(or_matrix, na.rm = TRUE), length.out = 100)
color_palette <- colorRampPalette(c("blue", "white", "red"))(length(breaks) - 1)

# Rename quantile columns to Q1, Q2, ..., Q25
colnames(or_matrix) <- paste0("Q", colnames(or_matrix))

# Plot heatmap
pheatmap(or_matrix,
         cluster_rows = TRUE,
         cluster_cols = FALSE,
         main = "OR by PRS Quantile (Ref: Q13)",
         fontsize_row = 10,
         fontsize_col = 10,
         color = color_palette,
         breaks = breaks,
         border_color = NA,
         angle_col = 45,  # Rotate quantile labels for readability
         legend_breaks = c(min(breaks), 1, max(breaks)),
         legend_labels = c("0.6", "OR = 1", "1.3"))
```

```{r}
#Volcano Plot
prev_df %>% split(., .$code) %>% lapply(function(mindf){
  q0 <- subset(mindf, PRS_quantile == 1) %>% select(cases, controls)
  qother <- subset(mindf, PRS_quantile != 1) %>% summarise_at(c('cases', 'controls'), sum)
  tt <- rbind(q0, qother) 
  rownames(tt) <- c('Q1', 'Other') 
  tt %>% as.matrix %>% fisher.test() -> tt
  return(c('pvalue' = tt$p.value, 'X2' = tt$estimate))

}) -> tt_2
tt_2 <- do.call('rbind', tt_2) %>% as.data.frame
tt_2$code <- rownames(tt_2)
colnames(tt_2)[2] <- 'OR'

tt_2$signif <- tt_2$pvalue <= 0.05
tt_2$adj.pvalue <- p.adjust(tt_2$pvalue)
tt_2$adj.signif <- tt_2$adj.pvalue <= 0.05
tt_2$sign.label <- ifelse(tt_2$adj.signif, '**', ifelse(tt_2$signif, '*', ''))

tt_2 <- tt_2 %>%
  mutate(
    chapter = str_sub(code, 1, 1),
    chapter_name = case_when(
      chapter %in% c("A", "B") ~ "Infectious and parasitic",
      chapter %in% c("C", "D") & as.numeric(str_sub(code, 2, 3)) <= 48 ~ "Neoplasms",
      chapter == "D" ~ "Blood and hematopoietic organs",
      chapter == "E" ~ "Endocrine, nutritional and metabolic",
      chapter == "F" ~ "Mental and behavioural disorders",
      chapter == "G" ~ "Nervous system",
      chapter == "H" & as.numeric(str_sub(code, 2, 3)) <= 59 ~ "Eye and adnexa",
      chapter == "H" ~ "Ear and mastoid process",
      chapter == "I" ~ "Circulatory system",
      chapter == "J" ~ "Respiratory system",
      chapter == "K" ~ "Digestive system",
      chapter == "L" ~ "Skin and subcutaneous tissue",
      chapter == "M" ~ "Musculoskeletal system and connective tissue",
      chapter == "N" ~ "Genitourinary system",
      chapter == "O" ~ "Pregnancy, childbirth and the puerperium",
      chapter == "P" ~ "Perinatal period",
      chapter == "Q" ~ "Congenital malformations",
      chapter == "R" ~ "Abnormal clinical and laboratory findings",
      chapter %in% c("S", "T") ~ "Injury and poisoning",
      chapter %in% c("V", "W", "X", "Y") ~ "External causes",
      chapter == "Z" ~ "Factors influencing health status",
      TRUE ~ "Codes for special purposes"
    )
  )


# Plot
ggplot(tt_2, aes(x = log2(OR), y = -log10(pvalue), color = chapter_name)) +
  
  # Points by ICD chapter
  geom_point(alpha = 0.8, size = 2, show.legend = TRUE) +

  # Horizontal line for p = 0.05
  geom_hline(yintercept = -log10(0.05), linetype = "dashed", color = "black") +

  # Vertical line for OR = 1 (log2(OR) = 0)
  geom_vline(xintercept = 0, linetype = "dashed", color = "darkred", linewidth = 0.8) +

  # Labels for adj.signif in black
  geom_text(data = subset(tt_2, adj.signif), 
            aes(label = code), 
            color = "black", 
            size = 4, 
            vjust = -0.5,
            show.legend = FALSE) +

  # Labels for signif but not adj.signif in light gray
  geom_text(data = subset(tt_2, signif & !adj.signif), 
            aes(label = code), 
            color = "gray", 
            size = 4, 
            vjust = -0.5,
            show.legend = FALSE) +

  # Axis labels and title
  labs(
    title = "Odds Ratio (Q1 vs Others)",
    x = "log2(Odds Ratio)",
    y = "-log10(p-value)",
    color = "ICD Chapter"
  ) +

  # Theme adjustments
  theme_minimal(base_size = 13) +
  theme(
    legend.position = "bottom",
    legend.title = element_text(size = 12),
    legend.text = element_text(size = 10),
    plot.title = element_text(hjust = 0.5, face = "bold"),
    panel.grid.major = element_line(color = "grey90"),
    panel.grid.minor = element_blank()
  )
```

```{r}
#Or table
selected_codes <- c("E11", "E66", "G56", "J43", "K59", "K62", "K92", "M17",
                    "M18", "M19", "M23", "M65", "N17", "R10", "Z37")

# Funci칩n para extraer OR, CI y p-valor del test de Fisher
get_fisher_summary <- function(code_input) {
  df_code <- df_long %>%
    filter(code == code_input) %>%
    mutate(quantil = if_else(PRS_quantile == 1, "Q1", "Others"))

  contingency <- df_code %>%
    group_by(quantil) %>%
    summarise(
      cases = sum(diagnosis == 1, na.rm = TRUE),
      controls = sum(diagnosis == 0, na.rm = TRUE),
      .groups = "drop"
    )

  # Build 2x2 matrix
  mat <- matrix(c(contingency$cases, contingency$controls), nrow = 2, byrow = TRUE)

  # Fisher test
  result <- tryCatch({
    ft <- fisher.test(mat)
    tibble(
      code = code_input,
      OR = round(unname(ft$estimate), 2),
      CI = paste0("[", round(ft$conf.int[1], 2), ", ", round(ft$conf.int[2], 2), "]"),
      p_value = signif(ft$p.value, 3)
    )
  }, error = function(e) {
    tibble(code = code_input, OR = NA, CI = NA, p_value = NA)
  })

  return(result)
}

# Aplicar a todos los c칩digos
summary_fisher <- map_df(selected_codes, get_fisher_summary)

# Mostrar
print(summary_fisher)
```