---
title: "Models"
subtitle: "Schizophrenia risk prediction_LMTV3"
author:
  - "Version: v1.0"
  - "Author: Pol Ramon-Cañellas"
  - "Updated by: Carla Arxé Asensio"
date: "Date: `r format(Sys.time(), '%Y.%m.%d')`"
output:
  html_document:
    toc: yes
    toc_depth: 4
    toc_float: yes
    toc_collapsed: yes
    code_folding: hide
    highlight: tango
    bs_theme: cerulea
  word_document:
    toc: yes
    toc_depth: 4
  pdf_document:
    toc: yes
    toc_depth: 4
---

# INTRODUCTION TO SCRIPT
SCRIPT PRUPOUSE: Check models for phenotype (Case/Control)(<https://bookdown.org/ndphillips/YaRrr/linear-regression-with-lm.html>)

- Approach 0:
  - 0.1: Binomial logistic regression model for phenotype (Case/Control), including covariates (PCs, Sex).
  - 0.2: Binomial logistic regression model for phenotype (Case/Control), including PCA values (PRS-PC1).

- Approach A: \n Binomial logistic regression model for phenotype (Case/Control), including covariates (PCs, Sex) and PCA values (PRS-PC1).
    
- Approach B:
  - B1. Linear regression model for PRS-PC1 to get the residual (E) using the covariates (PCs, Sex).
  - B2. Binomial logistic regression model for phenotype (Case/Control), using the PRS-PC1residual (E).

```{r starting, echo=FALSE}
# IMP: to free memory usage (when working with data requiring big amounts of memory)
rm(list=ls(all.names=TRUE))
invisible(gc())

#Libraries suppressMessages(library(dplyr))
suppressMessages(library(DescTools))
suppressMessages(library(dplyr))
suppressMessages(library(jtools))
suppressMessages(library(purrr))
suppressMessages(library(report))
suppressMessages(library(sjPlot))
suppressMessages(library(tidyr))

# INPUTS and data.frame rearrangements:

  # Phenotype (FID, IID, PHENOTYPE) 
  df_pheno <- read.table("1J_FID.IID.PHENOTYPE.txt", header=TRUE, sep=" ")
  df_pheno$FID_IID <- paste(df_pheno[,1], df_pheno[,2], sep="_|_")
  df_pheno <- df_pheno[ , c(length(df_pheno), 3:(length(df_pheno)-1))]  
  
    # Later, for models, the phenotype variable must be "0" (change "1" in controls) and "1" (change "2" in cases).
    #df_pheno[which(df_pheno$PHENOTYPE==1),"PHENOTYPE"] <- 0 
    #df_pheno[which(df_pheno$PHENOTYPE==2), "PHENOTYPE"] <- 1
  
  # Covariances (FID, IID, SEX, PC1,..., PC10) 
  df_cov <- read.table("1M_FID.IID.COVARIATES.txt", header=TRUE, sep=" ")
  df_cov$FID_IID <- paste(df_cov[,1], df_cov[,2], sep="_|_")
  df_cov <- df_cov[ , c(length(df_cov), 3:(length(df_cov)-1))]
  
  # PRS-PC1 (FID, IID, PRS.PC1)
  df_prs.pc1 <- read.table("3_PRS_individuals_PC1.txt", header=TRUE, sep=" ")
  names(df_prs.pc1)[names(df_prs.pc1)=="PC1"] <-"PRS.PC1"
  df_prs.pc1$FID_IID <- paste(df_prs.pc1[,1], df_prs.pc1[,2], sep="_|_") 
  df_prs.pc1 <- df_prs.pc1[ , c(length(df_prs.pc1), 3:(length(df_prs.pc1)-1))]
  
# MERGE ALL DATABASES
l_all <- list("pheno"=df_pheno, "covariates"=df_cov, "prs.pc1"=df_prs.pc1) 
df_all <- Reduce(function(x, y) merge(x, y, all=FALSE), l_all)
rownames(df_all) <- df_all$FID_IID

# PLOT DATA
par(mfrow=c(2,3))
for (i in seq_along(3:NCOL(df_all))) { 
  plot(df_all[,c(2,i+2)]) 
}
```

# APPROACH 0: \nBinomial logistic regression model for PHENOTYPE

## 0.1: Including covariances (NULL).
- Binary dependent variable: Phenotype (control '0' & case '1').
- Independent variables: SEX & PCs.

```{r aproach0.1, echo=FALSE}
# MODEL 0.1: covariates (null)
df_all_0 <- df_all
mod_0.1 <- glm(formula = PHENOTYPE ~ SEX+PC1+PC2+PC3+PC4+PC5+PC6+PC7+PC8+PC9+PC10, data=df_all_0, family="binomial")
mod_0.1_sum <- summary(mod_0.1)
print(mod_0.1_sum)
pseudor2_nagelkerke_0.1 <- PseudoR2(mod_0.1, which="Nagelkerke")
cat(paste0("\nThe pseudo R2 for Nagelkerke is ", pseudor2_nagelkerke_0.1, ".\n "))
mod_0.1_report <- report(mod_0.1)
print(mod_0.1_report)
```

## 0.2: Including covariances and PRS-PC1
- Binary dependent variable: Phenotype (control '0' & case '1').
- Independent variables: SEX, PCs, & PRS-PC1.

```{r aproach0.2, echo=FALSE}
# MODEL 0.2: PRS-PC1.
mod_0.2 <- glm(formula = PHENOTYPE ~ PRS.PC1, data=df_all_0, family="binomial")
mod_0.2_sum <- summary(mod_0.2)
print(mod_0.2_sum)
pseudor2_nagelkerke_0.2 <- PseudoR2(mod_0.2, which="Nagelkerke")
cat(paste0("\nThe pseudo R2 for Nagelkerke is ", pseudor2_nagelkerke_0.2, ".\n "))
mod_0.2_report <- report(mod_0.2)
print(mod_0.2_report)
```


# APPROACH A: \n Binomial logistic regression model ALL AT ONCE.

- Binary dependent variable: Phenotype (control '0' & case '1'). 
- Independent variables: SEX, PCs & PRS.PC1.

```{r aproachA1, echo=FALSE}
df_all_A <- df_all 
mod_A1 <- glm(formula = PHENOTYPE ~ SEX+PC1+PC2+PC3+PC4+PC5+PC6+PC7+PC8+PC9+PC10+PRS.PC1, data=df_all_A, family="binomial")
mod_A1_sum <- summary(mod_A1)
print(mod_A1_sum)
pseudor2_nagelkerke_A1 <- PseudoR2(mod_A1, which="Nagelkerke")
cat(paste0("\nThe pseudo R2 for Nagelkerke is ", pseudor2_nagelkerke_A1, ".\n "))
mod_A1_report <- report(mod_A1)
print(mod_A1_report)

#Check model fitting new predicted data. #Get the fitted values from the model and put them back in the dataset.
df_all_A$pred.PHENOTYPE <- mod_A1$fitted.values 
#Get the predictions of new data 
logit.predictions<- predict(object=mod_A1) 
#logit-transformed probabilities
prob.predictions <- 1 / (1 + exp(-logit.predictions)) 
# Apply inverse logit to transform to probabilities 
prob.predictions <- as.data.frame(prob.predictions)

#Plot original vs predicted values. 
mod_A1_pheno.pred <- merge(df_all_A[,'PHENOTYPE', drop=FALSE], prob.predictions, by=0, all=TRUE)

par(mfrow=c(2,2))

h0 <- (hist(mod_A1_pheno.pred[which(mod_A1_pheno.pred$PHENOTYPE==0),3], main="Control (0) prediction", xlab="Prediction"))
h0$density <- h0$counts/sum(h0$counts)*100
plot(h0,freq=FALSE, main="Control (0) prediction", ylab="Density (%)", xlab="Prediction")

h1 <- (hist(mod_A1_pheno.pred[which(mod_A1_pheno.pred$PHENOTYPE==1),3], main="Case (1) prediction", xlab="Prediction"))
h1$density <- h1$counts/sum(h1$counts)*100
plot(h1,freq=FALSE, main="Case (1) prediction", ylab="Density (%)", xlab="Prediction")
```

# APPROACH B: \n Linear regression model + Binomial logistic regresion model

## B1. Linear regression model for PRS-PC1:
- Dependent variable: PRS.PC1.
- Independent variables: SEX & PCs.

```{r aproachB1, echo=FALSE}
df_all_B <- df_all
mod_B1 <- lm(formula = PRS.PC1 ~ SEX+PC1+PC2+PC3+PC4+PC5+PC6+PC7+PC8+PC9+PC10, data=df_all_B)
mod_B1_sum <- summary(mod_B1) 
print(mod_B1_sum)
mod_B1_report <- report(mod_B1)
print(mod_B1_report)

cat(paste0("The covariables (SEX & PCs) explains the ", sprintf("%0.1f%%", mod_B1_sum[["adj.r.squared"]]*100), " of the PRS.PC1"))
cat(paste0("The residual (PRS.PC1.residual) explains the ", sprintf("%0.1f%%", (1-mod_B1_sum[["adj.r.squared"]])*100), " of the PRS.PC1"))

#Check model fitting new predicted data. #Get the fitted values from the model and put them back in the dataset.
df_all_B$pred.PHENOTYPE <- mod_B1$fitted.values 
#Get the predictions of new data 
logit.predictions<- predict(object=mod_B1) 
#logit-transformed probabilities 
prob.predictions <- 1 / (1 + exp(-logit.predictions)) 
#Apply inverse logit to transform to probabilities 
prob.predictions <- as.data.frame(prob.predictions)

#Plot original vs predicted values. 
mod_B_PRS.PC1.pred <- merge(df_all_B[,'PRS.PC1', drop=FALSE], prob.predictions, by=0, all=TRUE)

par(mfrow=c(1,1))

plot(mod_B_PRS.PC1.pred$PRS.PC1, mod_B_PRS.PC1.pred$prob.predictions, main="Prediction for PRS-PC1", xlab="PRS.PC1", ylab="prediction", xlim=c(min(mod_B_PRS.PC1.pred[,c(2,3)]), max(mod_B_PRS.PC1.pred[,c(2,3)])), ylim=c(min(mod_B_PRS.PC1.pred[,c(2,3)]), max(mod_B_PRS.PC1.pred[,c(2,3)])))

# Get the residuals PRS-PC1-res 
df_res <- as.data.frame(mod_B1$residuals)  
colnames(df_res) <- "PRS.PC1.residual"  
df_res$FID_IID <- rownames(df_res) 
df_res <- df_res[,c(2,1)]

#CONFIRM THAT PRS.PC1.residual IS NOT EXPLAINED BY COVARIABLES.
df_mod.res <- merge(df_all, df_res, by=c("FID_IID"))

mod_res <- lm(formula = PRS.PC1.residual ~ SEX+PC1+PC2+PC3+PC4+PC5+PC6+PC7+PC8+PC9+PC10, data=df_mod.res) 
mod_res_sum <- summary(mod_res) 
print(mod_res_sum)
cat(paste0("The residual (PRS.PC1.residual) is NOT explained by covariances, as expected."))
report(mod_res)
```

## B2. Binomial logistic regression model for phenotype (Case/Control):
- Binary dependent variable: Phenotype (control '0' & case '1').
- Independent variables: PRS.PC1.residual.

```{r aproachB2, echo=FALSE}
df_mod.B2 <- merge(df_pheno, df_res, by="FID_IID") 
rownames(df_mod.B2) <- df_mod.B2$FID_IID

mod_B2 <- glm(formula = PHENOTYPE ~ PRS.PC1.residual, data=df_mod.B2, family="binomial") 
mod_B2_sum <- summary(mod_B2)
print(mod_B2_sum)
pseudor2_nagelkerke_B2 <- PseudoR2(mod_B2, which="Nagelkerke")
cat(paste0("\nThe pseudo R2 for Nagelkerke is ", pseudor2_nagelkerke_B2, ".\n "))
mod_B2_report <- report(mod_B2)
print(mod_B2_report)

#Check model fitting new predicted data. #Get the fitted values from the model and put them back in the dataset.
df_mod.B2$pred.PHENOTYPE <- mod_B2$fitted.values 
#Get the predictions of new data 
logit.predictions <- predict(object=mod_B2) 
#logit-transformed probabilities 
prob.predictions <- 1 / (1 + exp(-logit.predictions)) 
# Apply inverse logit to transform to probabilities
prob.predictions <- as.data.frame(prob.predictions)

#Plot original vs predicted values. 
mod_B2_pheno.pred <- merge(df_mod.B2[,'PHENOTYPE', drop=FALSE], prob.predictions, by=0, all=TRUE)

par(mfrow=c(2,2))

h0 <- (hist(mod_B2_pheno.pred[which(mod_B2_pheno.pred$PHENOTYPE==0),3], main="Control prediction (0)", xlab="Prediction"))  
h0$density <- h0$counts/sum(h0$counts)*100 
plot(h0,freq=FALSE, main="Control prediction (0)", ylab="Density (%)", xlab="Prediction")

h1 <- (hist(mod_B2_pheno.pred[which(mod_B2_pheno.pred$PHENOTYPE==1),3], main="Case prediction (1)", xlab="Prediction"))  
h1$density <- h1$counts/sum(h1$counts)*100 
plot(h1,freq=FALSE, main="Case prediction (1)", ylab="Density (%)", xlab="Prediction")

```

# COMPARISON OF APPROACHES

Compare both approaches for the PHENOTYPE
```{r comparison, echo=FALSE}
models_sum <- sum(pseudor2_nagelkerke_B2, pseudor2_nagelkerke_0.1)
if (between(pseudor2_nagelkerke_A1, models_sum-0.03, models_sum+0.03)) {
  similar <- ""
} else {
  similar <- "NOT "
}
cat(paste0("The Nagelkerke pseudo R2 for the full model (A1) is ", pseudor2_nagelkerke_A1, ".\nWhich is ", similar, "similar to the sum of the PRS-PC1-residual model (B2), ", pseudor2_nagelkerke_B2, ", and the covariances model (0.1), ", pseudor2_nagelkerke_0.1, ", that sums ", models_sum, "."))


#Summary:
cat("#### SUMMARY OF MODELS ####")
export_summs(mod_0.1, mod_0.2, mod_A1, mod_B1, mod_B2, model.names=c("Approach 0.1", "Approach 0.2", "Approach A1", "Approach B1", "Approach B2"))

```

```{r save, echo=FALSE}
## Get from dataframe only PRS.PC1.residuals and separates FID_IID.
  df_res <- df_res %>% separate_wider_delim(FID_IID, "_|_", names=c("FID", "IID"))
  df_res <- as.data.frame(df_res)

#Save data in files
write.table(df_res, "3_PRS_PC1_res.txt", quote=FALSE, col.names=TRUE, row.names=FALSE)
```
